Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-05-16 paper/ms.tex
update ms

\appendix
\section{Cases With ``Inconsistent'' Rotation and Lithium-Based Ages}
\label{app:inconsistent}

>>>>

\appendix
\section{Literature Age Scale Comparison}
\label{app:litagecomparison}

Figure~\ref{fig:agescalecompone} shows the result of a literature comparison against
ages reported by \citet{Berger_2020a_catalog}, \citet{2021AJ....161..189L}, 
 \citet{2023ApJ...952..131M}, and
\citet{2024AJ....167..159L}.
We followed any guidance available from these respective studies for removing ages
that were flagged as unreliable.
Rather than compare the ages of the stars directly against each other, we then decided
to compare them against a common benchmark: open clusters.

Cluster memberships are a delicate subject.
Figure~\ref{fig:agescalecompone} is in fact showing putative ages for a set of putative cluster members.
For NGC6811 and NGC6819, we adopted all reported members from \citet{2018A&A...618A..93C,CantatGaudin_2020} and \citet{Kounkel_2020}. 
For Melange-3 we used the members from \citet{2022AJ....164...88B} and required ``offset'' tangential velocities below 2\kms.
For Cep-Her we used stars from Table~2 of \citet{Bouma_2022b}, requiring ``weights'' exceeding 0.1.
These cluster lists almost certainly include non-members, in fractions that we do not aim to quantify.
Even with these contaminants, we can compare the relative age distributions derived by
the different studies.

There are two main metrics for success in this test:
the first is the extent to which the reported ages agree with the cluster age (accuracy).
The second concerns how precise the reported ages are, and whether that precision agrees with the observed
distribution of ages around unity.

Starting with the isochrone ages from \citet{Berger_2020a_catalog}, the results should not be particularly surprising.
The isochrone ages are most accurate and precise (in relative terms) for the oldest cluster, NGC6819.
For the 1\,Gyr NGC6811, about half the stars have ages overestimated by a factor of three.
For the youngest clusters, the disagreement is greater.
The uncertainties seem accurate at 2.5\,Gyr, but perhaps less so at $\lesssim$1\,Gyr.
An important virtue of these isochrone ages is that they are calculated for far more stars in the two largest
clusters than the other methods.
They are calculable for far more stars than rotation-based ages.

The ``gyro-kinematic'' ages from \citet{2021AJ....161..189L} do...
{\bf todo explain}

The rotation-only ages calculated using the model described by \citet{2024AJ....167..159L} have some
peculiarities compared against cluster ages.
While NGC6811 seems to be in general agreement, NGC6819 has an abnormally large population of
young reported systems, with apparently precise ages.  All the younger cluster members
($<$300\,Myr) similarly have reported ages that are biased old, with narrow uncertainties.

\citet{2023ApJ...952..131M} looks great.  {\bf todo explain method}.
However it still fails for the youngest clusters.

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.98\textwidth]{merged_lit_ages_vs_cluster_ages.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Literature age scale comparison.}
    Each point denotes the reported age of a star observed by Kepler,
    normalized by the age of a stellar ensemble in which the star
    is a putative member.
    Rows show the available stellar ensembles, and
    columns show the various age-measurement techniques described in Appendix~\ref{app:litagecomparison}.
    Stars are sorted according to their median reported ages.
    Note that the horizontal axis is logarithmic; the first dotted line
    to the right of unity corresponds to a reported age ten times older
    than the cluster age.
	}
	\label{fig:agescalecompone}
\end{figure*}

\section{Cases With ``Inconsistent'' Rotation and Lithium-Based Ages}
\label{app:inconsistent}
